---
title: "Central Parking System"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

## Libraries...

```{r}
library(readxl)
library(skimr)
```

## Set path :-

```{r}
setwd("/Users/himanshusahrawat/MBA projects/CPS/R/")
```

## Read the excel data at sheet zero...

```{r}
data=read_excel("Central parking dataset.xls", sheet = 1)
data_s=read_excel("Central parking dataset.xls", sheet = 2)
```

## Removing unwanted coloumns...

```{r}
names(data)
print("--------------------------------------------------------------------------------")
names(data_s)
```

```{r}
data=data[ , c("Vehicle","Equipment","DateIn","Time In","DateOut","Time Out","Amount","TimeDiff","Ticket_Type","Weekday" )]
```

```{r}
data_s=data_s[ , c("Vehicle","Equipment","DateIn","...4","DateOut","...6","Amount","TimeDiff","Ticket_Type","Weekday" )]
```

## Renaming Columns for weekend data...

```{r}
colnames(data_s)[4] <- "Time In"
colnames(data_s)[6] <- "Time Out"
```

## Summary for weekday data...

```{r echo=TRUE, paged.print=TRUE}
skim(data)
```

## observations from weekday data :-

── Data Summary ──────────────────────── Values Name data\
Number of rows 5000\
Number of columns 10\
\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\
Column type frequency:\
character 2\
numeric 4\
POSIXct 4\
\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\
Group variables None

── Variable type: character ────────────────────────────────────────── skim_variable n_missing complete_rate min max empty n_unique 1 Ticket_Type 0 1 6 6 0 2 2 Weekday 0 1 6 9 0 5 whitespace 1 0 2 0

── Variable type: numeric ──────────────────────────────────────────── skim_variable n_missing complete_rate mean sd p0 p25 p50 p75 1 Vehicle 0 1 1 0 1 1 1 1 2 Equipment 0 1 5.63 0.913 5 5 5 6 3 Amount 0 1 36.8 28.5 30 30 30 40 4 TimeDiff 0 1 136. 83.9 2 76 122 179 p100 hist 1 1 ▁▁▇▁▁ 2 8 ▇▅▁▁▂ 3 300 ▇▁▁▁▁ 4 722 ▇▅▁▁▁

── Variable type: POSIXct ──────────────────────────────────────────── skim_variable n_missing complete_rate min\
1 DateIn 0 1 2009-07-06 00:00:00 2 Time In 0 1 1899-12-31 00:01:00 3 DateOut 0 1 2009-07-06 00:00:00 4 Time Out 0 1 1899-12-31 00:00:00 max median n_unique 1 2012-04-20 00:00:00 2010-12-28 00:00:00 696 2 1899-12-31 23:58:00 1899-12-31 16:24:00 880 3 2012-04-20 00:00:00 2010-12-28 00:00:00 760 4 1899-12-31 23:59:00 1899-12-31 18:11:00 952

## Summary for weekend data...

```{r}
skim(data_s)
```

## observation from weekend data ....

── Data Summary ──────────────────────── Values Name data_s Number of rows 4995\
Number of columns 10\
\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\
Column type frequency:\
character 2\
numeric 4\
POSIXct 4\
\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\
Group variables None

── Variable type: character ────────────────────────────────────────── skim_variable n_missing complete_rate min max empty n_unique 1 Ticket_Type 0 1 4 6 0 3 2 Weekday 0 1 6 8 0 2 whitespace 1 0 2 0

── Variable type: numeric ──────────────────────────────────────────── skim_variable n_missing complete_rate mean sd p0 p25 p50 p75 1 Vehicle 0 1 1 0 1 1 1 1 2 Equipment 0 1 5.95 1.13 5 5 6 6 3 Amount 0 1 62.5 31.4 0 50 50 70 4 TimeDiff 0 1 154. 87.7 1 91 141 200 p100 hist 1 1 ▁▁▇▁▁ 2 9 ▇▆▁▃▁ 3 300 ▇▃▁▁▁ 4 713 ▇▆▁▁▁

── Variable type: POSIXct ──────────────────────────────────────────── skim_variable n_missing complete_rate min\
1 DateIn 0 1 2009-07-04 00:00:00 2 Time In 0 1 1899-12-31 00:01:00 3 DateOut 0 1 2009-07-04 00:00:00 4 Time Out 0 1 1899-12-31 00:00:00 max median n_unique 1 2012-03-31 00:00:00 2010-10-30 00:00:00 279 2 1899-12-31 23:58:00 1899-12-31 16:39:00 852 3 2012-04-01 00:00:00 2010-10-30 00:00:00 364 4 1899-12-31 23:59:00 1899-12-31 18:35:00 893

## plots for length of stay

## weekdays

```{r echo=TRUE}
plot(x = data$TimeDiff, y = data$Amount,
    xlab = "Length of Stay",
    ylab = "Amount",
    main = "Weekday"
)
```

```{r echo=TRUE}
plot(x = data_s$TimeDiff, y = data_s$Amount,
    xlab = "Length of Stay",
    ylab = "Amount",
    main = "Weekend"
)
```

```{r}
library(tidyverse)
#remove the zero and 300 amount
data = data %>% filter(Amount <300)
```

```{r}
plot(x = data$TimeDiff, y = data$Amount,
    xlab = "Length of Stay",
    ylab = "Amount",
    main = "Weekday"
)
```

```{r}
data_s = data_s %>% filter(Amount <300) %>% filter(Amount>0)

plot(x = data_s$TimeDiff, y = data_s$Amount,
    xlab = "Length of Stay",
    ylab = "Amount",
    main = "Weekend"
)

```

```{r}
library(plotly)
```

Plot for count on daily basis.

```{r}
library(ggplot2)
```

```{r}
ggplot(data, aes(x = Weekday)) +
  geom_bar()
```

```{r}
ggplot(data_s, aes(x = Weekday)) +
  geom_bar()
```

Lets check the size of vehicles for both weekdays and weekends....

```{r}
data %>% filter(Amount>30 & TimeDiff<181)
```

```{r}
data_s %>% filter(Amount>60 & TimeDiff<181)
```

Adding colloumn for timing, if morning (00:00:00 - 12:00:00), afternoon (12:00:00 - 18:00:00), evening (18:00:00 - 00:00:00)

```{r}
# library(data.table)
 library(readxl)
# install.packages("skimr")
 library("skimr")
# install.packages("tidyverse")
 library("tidyverse")
# install.packages("plotly")
 library("plotly")
# install.packages("lubridate")
 library("lubridate")
# install.packages("recommenderlab")
 library("recommenderlab")
```

```{r}
class(data$`Time In`)
```

```{r}
#hour(data$`Time In`)
```

```{r}
# Multiple conditions when adding new column to dataframe:
data=data %>% mutate(Timing =
                     case_when((hour(`Time In`) >=00) & (hour(`Time In`)) <12 ~ "Morning", 
                               (hour(`Time In`) >=12) & (hour(`Time In`)) <18 ~ "Afternoon",
                               (hour(`Time In`) >=18) & (hour(`Time In`)) <24 ~ "Evening"))
```

```{r}
data_s=data_s %>% mutate(Timing =
                     case_when((hour(`Time In`) >=00) & (hour(`Time In`)) <12 ~ "Morning", 
                               (hour(`Time In`) >=12) & (hour(`Time In`)) <18 ~ "Afternoon",
                               (hour(`Time In`) >=18) & (hour(`Time In`)) <24 ~ "Evening"))
```

```{r}
tw = data %>% 
  group_by(Timing) %>%
  summarise(count =n())
  
```

```{r}
tw
```

```{r}

fig <- plot_ly(tw, labels = ~Timing, values = ~count, type = 'pie')

fig <- fig %>% layout(title = 'Weekday Timing',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
fig
```

people are coming in afternoon

```{r}
tw = data_s %>% 
  group_by(Timing) %>%
  summarise(count =n())


fig <- plot_ly(tw, labels = ~Timing, values = ~count, type = 'pie')

fig <- fig %>% layout(title = 'Weekend Timing',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
fig
```

During Saturday Sunday , most of the person are coming in afternoon.

Distribution of time spent by vehicles -----

```{r}

library(ggplot2)

```

```{r}
## Basic histogram from the vector "rating". Each bin is .5 wide.
## These both result in the same output:
ggplot(data, aes(x=TimeDiff)) + geom_histogram(binwidth=10)
# qplot(dat$rating, binwidth=.5)

# Draw with black outline, white fill
ggplot(data, aes(x=TimeDiff)) +
    geom_histogram(binwidth=10, colour="black", fill="white")

# Density curve
ggplot(data, aes(x=TimeDiff)) + geom_density()

# Histogram overlaid with kernel density curve
ggplot(data, aes(x=TimeDiff)) + 
    geom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=10,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")  # Overlay with transparent density plot
```

Checking Skew Value

```{r}
install.packages("moments")
library(moments)
```

```{r}
skewness(data$TimeDiff)
```

.If the skewness is between -0.5 and 0.5, the data are fairly symmetrical . If the skewness is between -1 and --- 0.5 or between 0.5 and 1, the data are moderately skewed . If the skewness is less than -1 or greater than 1, the data are highly skewed

DISTRIBUTION FOR WEEKEND

```{r}
## Basic histogram from the vector "rating". Each bin is .5 wide.
## These both result in the same output:
ggplot(data_s, aes(x=TimeDiff)) + geom_histogram(binwidth=10)
# qplot(dat$rating, binwidth=.5)

# Draw with black outline, white fill
ggplot(data_s, aes(x=TimeDiff)) +
    geom_histogram(binwidth=10, colour="black", fill="white")

# Density curve
ggplot(data_s, aes(x=TimeDiff)) + geom_density()

# Histogram overlaid with kernel density curve
ggplot(data_s, aes(x=TimeDiff)) + 
    geom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=10,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")  # Overlay with transparent density plot
```

```{r}
skewness(data_s$TimeDiff)
```

Q.4

People tends to stay longer during weekends.

```{r}
print("Average stay During weekday")
print(mean(data$TimeDiff))
```

```{r}
print("Average stay During weekend")
print(mean(data_s$TimeDiff))
```

Yes people tends to stay longer during weekends.

Q.5
```{r}
data=data %>% mutate(Timing =
                     case_when((hour(`Time In`) >=00) & (hour(`Time In`)) <14 ~ "Morning", 
                               (hour(`Time In`) >=14) & (hour(`Time In`)) <18 ~ "Afternoon",
                               (hour(`Time In`) >=18) & (hour(`Time In`)) <24 ~ "Evening"))
```


```{r}
tw = data %>% 
  group_by(Timing) %>%
  summarise(count =n())


fig <- plot_ly(tw, labels = ~Timing, values = ~count, type = 'pie')

fig <- fig %>% layout(title = 'Weekdays Timing',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
fig
```


```{r}

data_s=data_s %>% mutate(Timing =
                     case_when((hour(`Time In`) >=00) & (hour(`Time In`)) <14 ~ "Morning", 
                               (hour(`Time In`) >=14) & (hour(`Time In`)) <18 ~ "Afternoon",
                               (hour(`Time In`) >=18) & (hour(`Time In`)) <24 ~ "Evening"))
```

```{r}
tw = data_s %>% 
  group_by(Timing) %>%
  summarise(count =n())


fig <- plot_ly(tw, labels = ~Timing, values = ~count, type = 'pie')

fig <- fig %>% layout(title = 'Weekend Timing',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
fig
```

Q6.. If charges for first 2 hr is 20 and 10 for every additional hour... financial effect.

```{r}
new_price <- function(a) {
  b=as.integer(a[8])-120
  amount=20
  while (b>0) {
    amount=amount+10
    b=b-60
  }
  return(amount)
  }
```

```{r}
data$newprice = apply(data,1,new_price)
```

```{r}
sum(data$Amount)-sum(data$newprice)
```

loss of 24440


--------------------------------------------------------------------------------------------------------
q.8 analyst claim- aternate hypo- average ocupancy is not utmost 2 hr for weekdays. alpha - 0.05 . pop var=sam var


```{r}
data_sample=sample_n(data, 500)
```


```{r}

## Basic histogram from the vector "rating". Each bin is .5 wide.
## These both result in the same output:
ggplot(data_sample, aes(x=TimeDiff)) + geom_histogram(binwidth=.01)
# qplot(dat$rating, binwidth=.5)

# Draw with black outline, white fill
ggplot(data_sample, aes(x=TimeDiff)) +
    geom_histogram(binwidth=.01, colour="black", fill="white")

# Density curve
ggplot(data_sample, aes(x=TimeDiff)) + geom_density()

# Histogram overlaid with kernel density curve
ggplot(data_sample, aes(x=TimeDiff)) + 
    geom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=.01,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")
```

```{r}
install.packages("BSDA")
```

```{r}
library(BSDA)
```

h null : mean <=120  (claimed value)
h alter : mean >120 (opposite)

The alternative hypothesis in each case indicates the direction of divergence of the population mean for x (or difference of means for x and y) from mu (i.e., "greater", "less", "two.sided").

```{r}
data_sample=sample_n(data, 500)
z.test(x=data_sample$TimeDiff, mu=120, sigma.x=var(data$TimeDiff),alternative = "greater")
```

The test statistic for the one sample z-test is 0.054417 and the corresponding p-value is 0.4783.

Since this p-value is not less than .05, we have have sufficient evidence to reject the null hypothesis.



Thank you !!!!!
